Mathematical Aspects of the Capital Market

### 2017-07-21 phd/Chapterintroduction.tex
Added some paragraph in the introduction.

\section{Regular Variation}
The concept of regular variation is defined by the following scaling
property: If a function $f(\cdot)$ satisfies
\[
\lim_{x \to \infty} {f(c x) \over f(x)} = c^\alpha
\]
then we say $f(x)$ is regularly varying with index $\alpha$.
Distribution functions, say $F(\cdot)$, whose survival function
$\bar F(x) = 1 - F(x)$ satisfies the above scaling property, are
observed in a variety of phenomena, for example, phase transition of
matter, black body radiation, neuronal avalanches and stock returns.
When expanded to multiple dimensions, the scaling property is better
described in terms of weak convergence to a spectral measure: If a
random vector $V$ satisfies
\[
\lim_{x \to \infty}
{
  \P(V/|V| \in \cdot, |V| > c x)
  \over
  \P(|V| > x)
}
\overset{w}{\to} c^{-\alpha} \mu_\alpha(\cdot)
\]
then we say the components of $V$ is jointly regularly varying with
index $\alpha$. Here $\mu_\alpha(\cdot)$ is a probability measure on
the unit sphere \cite{buraczewski:damek:mikosch:2016}. Clearly, if $V$
is jointly regularly varying with index $\alpha$, then each component
of it and each linear combination of its components are regularly
varying with the same index $\alpha$. This follows from Feller
\cite{feller}, p. 278. Cf. also Jessen and Mikosch
\cite{JessenMikosch2006}, lemma 3.1, and Embrechts et al
\cite{embrechts:klueppelberg:mikosch:1997}, lemma 1.3.1.

@@

>>>>

\section{Regular Variation}
The concept of regular variation is defined by the following scaling
property: If a function $f(\cdot)$ satisfies
\[
\lim_{x \to \infty} {f(c x) \over f(x)} = c^\alpha
\]
then we say $f(x)$ is regularly varying with index $\alpha$.
Distribution functions, say $F(\cdot)$, whose survival function
$\bar F(x) = 1 - F(x)$ satisfies the above scaling property, are
observed in a variety of phenomena, for example, phase transition of
matter and black body radiation as studied in physics, neuronal
avalanches in biology, claim sizes of insurance mathematics and stock
returns in finance. The last application is indeed the focus of this
thesis.

When expanded to multiple dimensions, the scaling property of regular
variation is better described in terms of weak convergence to a
spectral measure: If a random vector $V$ satisfies
\[
\lim_{x \to \infty}
{
  \P(V/|V| \in \cdot, |V| > c x)
  \over
  \P(|V| > x)
}
\overset{w}{\to} c^{-\alpha} \mu_\alpha(\cdot)
\]
then we say the components of $V$ is jointly regularly varying with
index $\alpha$. Here $\mu_\alpha(\cdot)$ is a probability measure on
the unit sphere \cite{buraczewski:damek:mikosch:2016}. Clearly, if $V$
is jointly regularly varying with index $\alpha$, then each component
of it and each linear combination of its components are regularly
varying with the same index $\alpha$. This follows from Feller
\cite{feller}, p. 278. Cf. also Jessen and Mikosch
\cite{JessenMikosch2006}, lemma 3.1, and Embrechts et al
\cite{embrechts:klueppelberg:mikosch:1997}, lemma 1.3.1.

@@

When considered as the multiplicative inverse of the parameter of the
{\em Generalized Extreme Value} distribution, there are other methods
in the literature for estimating the tail index, e.g. Pickand's 
estimator \cite{pickands1975statistical} $\hat \alpha_P$ and the
Deckers-Einmahl-de Haan estimator $\hat \alpha_{\text{DEH}}$
\begin{eqnarray*}
{1 \over \hat \alpha_P}
&=&
{1 \over \log 2}
\log {
  X_{(k)} - X_{(2k)}
  \over
X_{(2k)} - X_{(4k)}  
} \\
{1 \over \hat \alpha_{\text{DEH}}}
&=&
1 + {1 \over \alpha_H} + {1 \over 2} \left(
  {H \over \alpha_H^2} - 1
\right)^{-1}
\end{eqnarray*}
where it is similarly assumed $k \to \infty$ and $k/n \to 0$ as $n \to
\infty$. The Deckers-Einmahl-de Haan estimator makes use of Hill's
estimator $\hat \alpha_H$ and computes
\[
H = \left[
  {1 \over k} \sum_{i=1}^k \log \left(
    { X_{(i)} \over X_{(k+1)} }
  \right)^2
\right]^{-1}
\]
Apparently $1/H$ can be interpreted as the 2nd empirical moment of
$\log(X_{(i)}/X_{(k+1)})$ for $i \leq k$.
A major drawback of Pickand's and Deckers-Einmahl-de Haan's estimators
is that, when applied to estimating the tail index, they discard the
information that the tail index is always positive, hence resulting in
a larger confidence band compared with that obtained for Hill's
estimator. Therefore we stick to Hill's estimator in the empirical
work included in this thesis.

One of the most important dynamical mechanisms that lead to regularly
varying r.v. is stochastic recursion of the following form:
\[
X_t = A_t X_{t-1} + B_t
\]
where $X_t$ is a $d$-dimensional random vector, $A_t$ is a $d\times d$
random matrix and $B_t$ is a $d$-dimensional vector, random or
deterministic.

\section{Stochastic Recurrence Equation}
